---
title: "7-DBSCAN"
output:
  html_document:
    df_print: paged
    toc: true
  html_notebook:
    highlight: textmate
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```

***

# dbscan Example

```{r}
# install.packages("dbscan") # install if necessary
library("dbscan") 
#install.packages("fpc")
library("fpc")
library('factoextra')
#moons dataset Contains 100 2-D points, half of which are contained in two # moons or “blobs” (25 points each blob), and the other half in asymmetric # facing crescent shapes.
data("moons")  
plot(moons, pch=20) # plot original data points

```

The data shows both nonlinearity and different densities. 

***

Let us apply dbscan to this example
```{r}
db <- dbscan::dbscan(moons, eps = 0.5, minPts = 5) #perform clustering

print(db) #print cluster details

plot(db, moons, main = "DBSCAN", frame = FALSE) #plot cluster details
```

***

# fpc Package

Let us now apply the fpc package to an example
```{r}
# install_packages("fpc")
library('factoextra')
library('fpc')
df <- multishapes[, 1:2]

set.seed(123)
db <- fpc::dbscan(df, eps = 0.15, MinPts = 5) # DBSCAN using fpc package

print(db) # show clusters' details

```

The results show that there are a total of 31 border points. The values under the row Seed indicate the core points. 

```{r}
# Plot DBSCAN results
plot(db, df, main = "DBSCAN", frame = FALSE)
```

The colored triangles indicate core points, while the colored circles border points. The black circles indicates noise (outliers).

Here is an alternative way to depict the same plot
```{r}
fviz_cluster(db, df,   stand = FALSE, frame = FALSE, geom = "point") # Alternative way to depict plot
```

***

## Optimal eps value

dbscan is sensitive to changes in eps values, as the below example illustrates.
```{r}
db <- fpc::dbscan(df, eps = 0.5, MinPts = 5)
plot(db, df, main = "DBSCAN", frame = FALSE)
```

As you can clearly see, changing the neighborhood changes the categorization of the points. Many of the noise points now belong to a cluster, and there are fewer border points.

***

To determine the optimal eps, we compute the the k-nearest neighbor distances in a matrix of points.The idea is to calculate, the average of the distances of every point to its k nearest neighbors. The value of k will be specified by the user and corresponds to MinPts. Next, these k-distances are plotted in an ascending order. The aim is to determine the “knee”, which corresponds to the optimal eps parameter.


```{r}
dbscan::kNNdistplot(df, k =  5)
abline(h = 0.15, lty = 2)
```

Based on the "knee" plot above, the optimal eps = 0.15. 
