---
title: 'Naive Bayes: Predicting Delayed Flights'
author: "Murali Shanker"
date: "10/23/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Problem Statement

Predicting flight delays can be useful to a variety of organizations. Here, we look at five predictors. The outcome of interest is whether or not the flight is delayed (delayed here means
arrived more than 15 minutes late). Our data consist of all flights from the Washington, DC area into the New York City area during January 2004. A record is a particular flight. The percentage of delayed flights among these 2201 flights is 19.5%. The data were obtained from the Bureau of Transportation Statistics (available on the web at www.transtats.bts.gov). The goal is to accurately predict whether or not a new flight (not in this dataset), will be delayed. The outcome variable is whether the flight was delayed, and thus it has two classes (1 = delayed and 0 = on time).

## Data Preparation

Load required libraries, and read data
```{r}
require(e1071)
require(caret) # use this to create datasets

delay <- read.csv("./FlightDelays.csv")
str(delay)
head(delay)
```
We need to convert DAY_WEEK and DAY_OF_MONTH into factors. We also need to categorize CRS_DEP_TIME. There are several ways of doing the last. Here, we will group time into buckets of one hour.
```{r}
delay$DAY_WEEK <- factor(delay$DAY_WEEK)
delay$DAY_OF_MONTH <- factor(delay$DAY_OF_MONTH)
delay$Flight.Status <- factor(delay$Flight.Status)

# Now put departure time into one-hour buckets

delay$CRS_DEP_TIME <- factor(round(delay$CRS_DEP_TIME/100))
str(delay)
head(delay)
```
***

## Questions

1. Divide the data into 60% training and 40% validation

```{r}
# First select the required variables
selected.var <- c(1, 2, 4, 8, 10, 13)
set.seed(15)
delay.tr.in = createDataPartition(delay$Flight.Status,p=0.6, list=FALSE) # 60% reserved for Training
delay.tr = delay[delay.tr.in,selected.var]
delay.va <- delay[-delay.tr.in, selected.var] # Validation  data is rest
summary(delay.tr)
summary(delay.va)
```

***
2. Output both a counts table and a proportion table outlining how many and what proportion of flights were delayed and on-time at each of the three airports.

We will use prop.table() to output the proportions from counts
```{r}
table(delay.tr$Flight.Status,delay.tr$DEST)
prop.table(table(delay.tr$Flight.Status,delay.tr$DAY_WEEK),margin = 1)
prop.table(table(delay.tr$Flight.Status,delay.tr$CRS_DEP_TIME),margin = 1)
prop.table(table(delay.tr$Flight.Status,delay.tr$CARRIER),margin = 1)
prop.table(table(delay.tr$Flight.Status,delay.tr$DEST),margin = 1)
prop.table(table(delay.tr$Flight.Status,delay.tr$ORIGIN),margin = 1)

 # This will sum row probabilities to 1
```

We can also the pivot table to generate proportions and counts
```{r}
attach(delay.tr)
ftable(Flight.Status,CARRIER,DEST)
detach(delay.tr)

# Or

prop.table(ftable(delay.tr, row.vars = 6, col.vars = 2), margin = 1)

```

****
3. Run the Naive Bayes model to predict whether the flight is delayed or not. 
```{r}
delay.nb <- naiveBayes(Flight.Status ~ ., data = delay.tr)
delay.nb
```
```{r}

```


***
4. Output the confusion matrix and ROC for the validation data

Let us output the confusion matrices for the training and validation sets, and determine the prediction for a (new) test set

```{r}
# Training set
pred.class <- predict(delay.nb, newdata = delay.tr)
confusionMatrix(pred.class, delay.tr$Flight.Status)

# validation
pred.prob <- predict(delay.nb, newdata=delay.va, type="raw") # probabilities
pred.class <- predict(delay.nb, newdata = delay.va) # class membership
confusionMatrix(pred.class, delay.va$Flight.Status)

# For the test set
df <- data.frame(actual = delay.va$Flight.Status, predicted = pred.class, pred.prob)

df[delay.va$CARRIER == "DL" & delay.va$DAY_WEEK == 7 & delay.va$CRS_DEP_TIME == 10 & 
     delay.va$DEST == "LGA" & delay.va$ORIGIN == "DCA",]

```

***

We will now output the ROC


```{r}
# install.packages("pROC") # install if necessary
require(pROC)

# Note the delayed probabilities are in column 1

roc(delay.va$Flight.Status,pred.prob[,1])
plot.roc(delay.va$Flight.Status,pred.prob[,1])

# Most confusion matrices use a threshold of 0.5. We can also output the threshold value that maximizes the sum of Sensitivity and Specificity, as shown below

plot.roc(delay.va$Flight.Status,pred.prob[,1],print.thres="best")

```
The output shows that using a cutoff of 0.146 produces the maximum value for Sensitivity (of 0.484) + Specificity (of 0.778).

For further reading on dealing with probability thresholds:

1. [Dealing with imbalanced classes](https://topepo.github.io/caret/using-your-own-model-in-train.html#Illustration5)
2. [A book dealing with optimizing probability thresholds](http://appliedpredictivemodeling.com/blog/2014/2/1/lw6har9oewknvus176q4o41alqw2ow)


```{r}
set.seed(949)
mod0 <- train(Flight.Status ~ ., data = delay.tr,
              method = "rf",
              metric = "ROC",
              tuneGrid = data.frame(mtry = 3),
              ntree = 1000,
              trControl = trainControl(method = "repeatedcv",
                                       repeats = 2,
                                       classProbs = TRUE,
                                       summaryFunction = twoClassSummary))
getTrainPerf(mod0)

roc0 <- roc(delay.va$Flight.Status, 
            predict(mod0, delay.va, type = "prob")[,1], 
            levels = rev(levels(delay.va$Flight.Status)))
roc0
plot(roc0, print.thres = c(.5), type = "S",
     print.thres.pattern = "%.3f (Spec = %.2f, Sens = %.2f)",
     print.thres.cex = .8, 
     legacy.axes = TRUE)
```


